---
title: "2021101113_Visualisation_Assignment"
author: "Gowlapalli Rohit"
date: "23/1/2024"
output:
  pdf_document:
    toc: true
  word_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

### Question 1
```{r}
# Load necessary libraries
library(stats)
library(graphics)

# Set seed for reproducibility
set.seed(123)

# Function to perform the required tasks
sampling_distribution_analysis <- function(sample_size, num_trials) {
  
  # Initialize vectors to store means and standard deviations
  means <- numeric(num_trials)
  std_devs <- numeric(num_trials)
  
  # Loop through the trials
  for (i in 1:num_trials) {
    
    # Generate a random sample from N(100,15) 
    sample_data <- rnorm(sample_size, mean = 100, sd = 15)
    
    # Calculate mean and standard deviation of the sample
    means[i] <- mean(sample_data)
    std_devs[i] <- sd(sample_data)
  }
  
  # Plot the frequency distribution of means
  hist(means, main = paste("Sampling Distribution (n =", sample_size, ")"),
       xlab = "Sample Mean", col = "skyblue", border = "black", prob = TRUE)
  
  # Add a normal curve to the plot for comparison
  curve(dnorm(x, mean = 100, sd = 15/sqrt(sample_size)), add = TRUE, col = "red", lwd = 2)
  
  # Plot the frequency distribution of standard deviations
  hist(std_devs, main = paste("Sampling Distribution of SD (n =", sample_size, ")"),
       xlab = "Sample Standard Deviation", col = "lightgreen", border = "black", prob = TRUE)
  
  # Add a normal curve to the plot for comparison
  curve(sqrt((15^2)/(sample_size-1)) * dchisq(x^2, df = sample_size-1), add = TRUE, col = "blue", lwd = 2)
}

# Perform the analysis for different sample sizes
sampling_distribution_analysis(10, 1000)
sampling_distribution_analysis(50, 1000)
sampling_distribution_analysis(100, 1000)
sampling_distribution_analysis(500, 1000)
sampling_distribution_analysis(1500, 1000)
```
In the context of the Central Limit Theorem (CLT), which states that the distribution of the sum (or average) of a large number of independent, identically distributed random variables approaches a normal distribution regardless of the original distribution, here are some observations and inferences from the produced histograms:

As Sample Size Increases, Distribution Approaches Normality:

As the sample size increases, you observe that the distribution of sample means becomes more bell-shaped and approaches a normal distribution.
This aligns with the central limit theorem, which suggests that the sampling distribution of the sample mean becomes approximately normal, even if the underlying population distribution is not normal.
Consistency of Sample Mean:

The mean of each sampling distribution should be close to the population mean (100, as specified in the problem).
This demonstrates the unbiased nature of the sample mean as an estimator of the population mean.
Reduced Variability with Larger Samples:

With larger sample sizes, the variability (standard deviation) of the sample means decreases.
This reduction in variability is in line with the central limit theorem, which states that larger sample sizes lead to a more concentrated sampling distribution around the true population mean.
Comparison with Theoretical Distributions:

The red curve in the histograms represents the theoretical normal distribution for the mean, and the blue curve represents the theoretical chi-square distribution for the standard deviation.
The close alignment of the histograms with these curves supports the central limit theorem's prediction that, with a sufficiently large sample size, the sampling distribution of the mean and standard deviation becomes normal.
Accuracy of Estimates:

As the sample size increases, the accuracy of estimating the population parameters (mean and standard deviation) improves.
This reflects the efficiency of larger sample sizes in providing more reliable estimates of population parameters.
Understanding the Role of Sample Size:

Smaller sample sizes may exhibit more variability, but as the sample size increases, the distribution becomes more centered around the true population mean.
In summary, the produced histograms and their alignment with theoretical distributions confirm the central limit theorem's predictions. Larger sample sizes lead to more normal-like sampling distributions of the mean and reduced variability, supporting the idea that larger samples provide more reliable estimates of population parameters.

