---
title: "Divyansh Tiwari (2020111002)"
output:
  pdf_document: default
  word_document: default
date: "2023-02-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

The following is the code for the 1st question.

```{r}

p_values <- c(0.0050, 0.0010, 0.0100, 0.0005, 0.0009, 0.0400, 0.0560, 0.0500, 0.0480, 0.0130, 0.0370, 0.0430, 0.0020, 0.0250, 0.1100, 0.0700, 0.0800)

p_values <- sort(p_values)

p_bonferroni <- p.adjust(p_values, method = "bonferroni", n = length(p_values))

p_hochberg <- p.adjust(p_values, method = "hochberg", n = length(p_values))


print(p_values)
print(p_bonferroni)
print(p_hochberg)

```


```{r}
library(reshape2)
library(ggplot2)

data <- data.frame(x = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17), y1 = p_values, y2 = p_bonferroni, y3 = p_hochberg, y4 = c(0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500, 0.0500))

gfg_plot <- ggplot(data, aes(x)) +  
    geom_line(aes(y = y1), color = "blue") +
     geom_line(aes(y = y2), color = "red") +
    geom_line(aes(y = y3), color = "green")+
    geom_line(aes(y = y4), color = "black", linetype = "twodash")+ labs(title="Question 1", x = "sorted outcomes", y = "p-values")

gfg_plot
#+geom_line(aes(color=c("Unadjusted", "Bonferroni", "Hochberg")))


```

The following is the inference from the graph: 

High adjusted p values for the earlier experiments show that there is not enough data to conclusively demonstrate an effect. The more severe Bonferroni correction produces bigger adjusted p values. It is less rigorous since the Hochberg adjustment correlates to lower adjusted p values for the experiment than our alpha. The bonferroni test is a conservative test since it has a high probability of rejecting the null hypothesis.

# Question 2

The following is the code for the second question.

## Part A

```{r}
library(readxl)
mydataq2 <- read_excel("./BRSM_Results_Visualization.xlsx", 1)

print(mydataq2)
beer <- mydataq2[mydataq2$Group == "Beer", ]
print(beer)
water <- mydataq2[mydataq2$Group == "Water", ]
print(water$`No. of Mosquitoes`)

beer_mean <- mean(beer$`No. of Mosquitoes`)
print(beer_mean)
water_mean <- mean(water$`No. of Mosquitoes`)
print(water_mean)
test.stat.1 <- abs(water_mean - beer_mean)

beer_median <- median(beer$`No. of Mosquitoes`)
water_median <- median(water$`No. of Mosquitoes`)
test.stat.2 <- abs(beer_median - water_median)
print(test.stat.2)
cat("The difference between the group medians is", test.stat.2)

```


```{r}
set.seed(1979)

n <- length(mydataq2$Group)
P <- 10000
var <- mydataq2$`No. of Mosquitoes`

PermSamples <- matrix(0, nrow = n, ncol = P)

for(i in 1:P)
{
  PermSamples[, i] <- sample(var, size = n, replace = FALSE)
}


Perm.test.stat1 <- rep(0, P)
Perm.test.stat2 <- rep(0, P)

for (i in 1:P)
{
    Perm.test.stat1[i] <- mean(PermSamples[mydataq2$Group == "Beer", i]) - mean(PermSamples[mydataq2$Group == "Water", i])
  
    Perm.test.stat2[i] <- median(PermSamples[mydataq2$Group == "Beer", i]) - median(PermSamples[mydataq2$Group == "Water", i])
  
}

print(mean(Perm.test.stat2 >= test.stat.2))

hist(Perm.test.stat2, col = "green", main="Distribution of median differences from 10k random permutations")
abline(v=test.stat.2, lwd=3, col="red")


```


```{r}

p_value = sum(Perm.test.stat2 >= test.stat.2)/P
print(p_value)
cat("The p_value obtained is ", p_value)

```

We observe that the p-value is less than 0.05 which brings to us a very high possibility of the alternate directional hypothesis to be true and the observed statistic is significant.

## Part B

```{r}

t <- t.test(beer$`No. of Mosquitoes`, water$`No. of Mosquitoes`, var.equal = TRUE)
t <- t$statistic[['t']]
cat("The initial t-score is ", t)
```

```{r}

p2_perm = rep(0, P)

for(i in 1:P)
{
  temp_t <- t.test(head(PermSamples[, i], 25), tail(PermSamples[,i], 18), var.equal = TRUE)
  p2_perm[i] <- temp_t$statistic[['t']]
}

p_value_2 = (sum(p2_perm >= t))/P
cat("The observed p_value is ", p_value_2)

```

The observed p_value is less than 0.05 which confirms that the alternate hypothesis is true and the observed statistic is significant.

```{r}

hist(Perm.test.stat2, col="green", main = "Distribution of t-scores from 10k random permutations")
abline(v=t, lwd=2, col="red")
```


## Part C

Since, we are calculating the p-value assuming that the hypothesis is non-directional, therefore we need to calculate the p-value with the absolute value of the differences of the medians.

```{r}

p_value_3 = sum(abs(Perm.test.stat2) >= test.stat.2)/P
cat("The p-value is this case is ", p_value_3)

```

The p-value is observed to be less than 0.05 which implies that the non-directional hypothesis is true and the observed statistic is significant.

# Question 3

Following is the code for the third question.

```{r}

mydataq3 <- read_excel("./IQ.xlsx", 1)
q3_p <- cor(mydataq3$IQ, mydataq3$TESTSCORE)
print("The intial correlation value is ")
print(q3_p)

n <- length(mydataq3$IQ)
P <- 10000
iq <- mydataq3$IQ
test_scores <- mydataq3$TESTSCORE

iq_samples <- matrix(0, nrow = n, ncol = P)
test_scores_samples <- matrix(0, nrow = n, ncol = P)

for(i in 1:P)
{
  iq_samples[, i] <- sample(iq, size = n, replace = FALSE)
  test_scores_samples[, i] <- sample(test_scores, size = n, replace = FALSE)
}

p3_perm = rep(0, P)

for(i in 1:P)
{
  
  p3_perm[i] = cor(iq_samples[, i], test_scores_samples[, i])
}
 
p_value_4 <- sum(p3_perm >= q3_p)/P
print(p_value_4)
print(paste("The observed p-value is ", p_value_4))

```
The p-value that is being obtained as I run the code above oscillates between 0 and 1e-04. At the very moment, the p-value from the code above is 1e-04, but for some reason it is getting rendered as 0 in the pdf.

We observe that the p-value is much less than 0.05 based on our observed statistic. Therefore, we can safely assume that we can reject the null hypothesis, since there exists a correlation between IQ and Test scores.

```{r}

hist(p3_perm, col="green", main = "Distribution of correlation values from 10k random permutations", xlim = c(-0.7, 0.7))
abline(v=q3_p, lwd=2, col="red")
```
